Notes on Semi-Supervised `Novelty Detection'

### 2011-04-09 semisupervised/notes.tex
Added experience 2 with dora world model.
In it, a class dependent and and a class independent probability distribution is learned on "known" density.
And a simple not-class-aware is used to model the unknown.

This is interesting as its makes it possible to model using a graphical models.

TODO: distributions are very strickyly calculated atm and they return 0 for values never seen.
That should be smoothed out.
Updates

The system objective was to detect novel data from a given known set of 6 categories.
Two tests where run:
\begin{enumerate}
\item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{tab:dora-data-perfect}).
\item A simple model was used to learn the probability functions and the thresholds were compared (\autoref{tab:dora-data-simple}).
\end{enumerate}

@@

\begin{figure}
\centering
\include{dora-data/roc-simple.tex}
\caption{\label{tab:dora-data-simple}ROC curve for thresholds learned on samples from samples in the dora distribution}
\end{figure}

\begin{sidewaystable}[h]
\begin{center}
\scalebox{0.35}{
\input{dora-data/explain}
}
\end{center}
\caption{\label{tab:dora-world-distribution}World model used in Dora. Each column cell shows $P(feature|class)$}
\end{sidewaystable}

>>>>

The system objective was to detect novel data from a given known set of 6 categories.
Two tests where run:
\begin{enumerate}
\item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{tab:dora-data-perfect}).
\item The models described in \autoref{sec:dora-models} were used to learn the probability functions and the thresholds were compared (\autoref{tab:dora-data-simple}).
\end{enumerate}

\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
      to the probabilities functions to not return 0. This is because at the moment if on the known
      samples we never see a given value we don't know it belongs to the possible feature space.
      Has to be corrected.}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{dora-data/roc-simple.pdf}
\caption{\label{tab:dora-data-simple}ROC curve for thresholds learned on samples from samples in the dora distribution}
\end{figure}

\begin{sidewaystable}[h]
\begin{center}
\scalebox{0.35}{\input{dora-data/explain}}
\end{center}
\caption{\label{tab:dora-world-distribution}World model used in Dora. Each column cell shows $P(feature|class)$}
\end{sidewaystable}

\subsection{How Dora models the world?}
\label{sec:dora-models}

Currently dora uses a \emph{probabilistic graphical model} to model the world.
All variables that the system see are added and connected in between with the
\emph{probabilistic factors} that the system is aware of.
Those factors are learned from the labelled data that the system is given when
training.

The proposal here is to also use such an approach to model the unconditional
probability.
Under that a novelty detection threshold can be easily implemented
reusing the \emph{probabilistic graphical model} $G$ that dora uses as
representation of the world:
\begin{description}
\item[$P(x|know)$] is obtained evaluating $P(G)$.
\item[$P(x)$] can be obtained by creating $G'$ as an instance of $G$ but with the
variable node we want to perform novelty detection removed and the factors associated
with it replaced with factors learned from the unlabelled data.
\end{description}

We point to an additional simplification that if the original graphical model factors
depending on the variable to perform novelty detection were previously split on a
conditional and unconditional factors. Then $G'$ could be obtained by simply removing
all the factors associated with the variable to perform novelty detection on.
